\Sof{TBA - New Titles in the comments}

### 2020-04-15 paper/icsme20.tex
paper update

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}
\usepackage{footnote}
\usepackage{xfrac}
\usepackage{balance}
\usepackage{url}

@@

\title{Fixing Vulnerabilities Hinders Maintainability}

@@

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of $10$ 
guidelines using a dataset including $1037$ security commits. Results show 
that fixing software vulnerabilities hinders the maintainability of 
open-source software. This is particularly the case for patches that address Path Traversal (CWE-22),
Information Exposure (CWE-200) and Resource Management Errors (CWE-399). 
We conclude that changes to code bases to fix vulnerabilities need to be 
performed with extra care and that security-related code evolution should 
be subject in computer curricula.
\end{abstract}

@@

exploit their resources. \Sof{Improve example - outdated}As an example, in the last week of January $2019$, 

@@

details available at \url{https://thehackernews.com/2019/01/apple-facetime-
privacy-hack.html} (Accessed on \today{})} --- the outcome of a design/logical
flaw that allowed callers to see and hear others without them picking up the call.%\cite{cacheOut}.

@@

maintainability~\cite{4724577} and documentation to help developers adopt security 
best practices~\cite{6311252, 7927935, MESQUIDA201519}.

@@

  \textit{<anonymized for double blind review>}.

@@

implications; section~\ref{sec:threats}, enumerates the threats to the validity of

@@

%%%%% NEEDS REVIEW
As an example, consider the patch of the TLS state-machine protocol implementation in OpenSSL\footnoteref{openssl} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting TLS messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2016$-$6307$ \footnote{CVE-$2016$-$6307$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2016-6307}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
evaluation. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/statem/statem.c} file\footnote{CVE-$2016$-$6307$ fix available  at
\url{https://github.com/openssl/openssl/commit/c1ef7c971d0bbf117c3c80f65b5875e2e7b024b1}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between the two parties. The problem in
CVE-$2016$-$6307$ is that due to a flaw in the logic of version $1.1.0$,
memory for a message is allocated too early, prior to the excessive
message length check which can lead to a Denial of Service through memory
exhaustion if an attacker triggers up a certain message length to be allocated to service a
connection. In this case, there is only a security impact if the application does not call  \texttt{SSL\_free()} in a timely manner in the
event that the connection fails;
or, the application is working in a constrained environment where there
is very little free memory;
or, the attacker initiates multiple connection attempts such that there
are multiple connections in a state where memory has been allocated for
the connection, \texttt{SSL\_free()} has not yet been called, and there is
insufficient memory to service the multiple requests. Except in the instance of $1$ above, any Denial Of Service is likely to
be transitory because as soon as the connection fails the memory is
subsequently freed again in the \texttt{SSL\_free()} call. However there is an
increased risk during this period of application crashes due to the lack
of memory - which would then mean a more serious Denial of Service.

@@

%
Although checking the maximum size of a message seems an elementary
problem to repair, the example shows that there was a significant amount of
changes performed in the code base which produced a negative impact in the
maintainability. The developer tries to patch the
code base by introducing a new branch point to the function \texttt{read\_state\_machine}~\ref{lst:func1} which adds complexity to the software
and disrupts
one of the guidelines proposed by the Software Improvement Group (SIG) for
building maintainable software~\cite{Visser:2016:OREILLY} $-$
the \emph{Write Simple Units of Code} guideline $-$ which resulted in a negative impact on the
OpenSSL maintainability.
%%%%%%% 
Software maintainability is designed as the degree to which an application is understood, repaired or enhanced. In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability of their
software applications. This is important because software maintainability is approximately $75\%$ of the cost related to a project. To answer the following three research questions, we use two datasets
of security patches~\cite{Reis:2017:IJSSE, 10.1109/MSR.2019.00064} to measure the impact of
security patches on the maintainability of open-source software. 
%

Often security flaws require patching code to make software more secure.
However, \textbf{there is no evidence yet of how security patches impact the
maintainability of open-source software}. Our suspicion is that developers tend
to introduce technical debt in their software when patching the code base to
address a security flaw because they tend to choose the easiest path to solve
it. To address it, we compute the maintainability of $1037$ commits using the
\emph{Better Code Hub} tool. 
%
\subsubsection*{\textit{\textbf{RQ2}  \textbf{Which patterns of security patches are more likely to
affect open-source software maintainability?}}}

@@

the cross-site scripting vulnerability is presented in Listing~\ref{lst:fix}. The developer added the function \texttt{htmlentities} to \textit{escape} the data given by the variable

@@

\subsubsection*{\textit{\textbf{RQ3} \textbf{What is the impact of security patches versus regular changes on the
maintainability of open-source software?}}}

@@

We also compute the maintainability of $1037$ random regular commits using the

@@

	Extract relevant data (e.g., sha, sha of the parent commit, owner of the project, name of 
	the repository) and code bases from the combined dataset containing $1335$ security

@@

the outcome of mining/manually inspect $267$ GitHub projects. Reis and Abreu 

@@

dataset of $711$ test cases for $16$ security patterns. Some of the patterns

@@

code used to fix the security flaw. In this study, we focus on computing the

@@

for references to commits involved in the patch on the NVD's website. However, $70\%$

@@

used to patch the vulnerability. We assumed that the last commit of the patch as 

@@

@@

$1037$ security patches. $556$ of these patches are identified with Common Weakness
Enumeration IDs (CWE). In Table~\ref{tab:patterns}, the most prevailing categories are presented. Patterns
with less than $20$ instances were merged in a bigger pattern named \textit{Miscellaneous}.

\begin{table}[h]
\footnotesize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.1cm}L{2.1cm}L{4.3cm}}

@@

 	     	           \textbf{CWE-79} & 
Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)& 
Lack of proper validation or escaping
 	    allows attackers to submit untrusted data to web browsers through malicious
 	    scripts that can hijack the user sessions or redirect the user to malicious
 	    sites.\\\midrule

@@

 	 	     	           \textbf{MISC} & 
Miscellaneous (673 cases) & 
This pattern comprises
 		several other security refactorings that do not have a CWE assigned and patterns that
 		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.\\
\bottomrule
\end{tabular}
\label{tab:patterns}
\end{table}

@@

The baseline dataset uses the security commits dataset as input.
Both datasets have the same size. For each
security commit, a commit with an approximate size is searched. Due to the complexity 
of some patches, it was not possible to find patches with the exact number of added and 
deleted lines. Our search uses the patch \emph{diff} to look for a new similar \emph{diff}. The algorithm follows the following steps:

\begin{enumerate}
\item Diff comparison between the security patch commit (\emph{sha}) and its parent (\emph{sha-p}).
\item Selects a random commit from the security patch project (\emph{sha-reg}).
\item Diff comparison between the random commit (\emph{sha-reg}) and its parent (\emph{sha-reg-p}).
\item Check if the size from the regular change has the same/approximated size to 
the security patch.
\end{enumerate}

Our approximation was calculated using an initial step of $1$. Every $10$ attempts to search for 
a change with the same \emph{diff} the step increased. Thus, if a change suffered $20$ deletions 
and $10$ additions, then, in the first $10$ attempts, the algorithm searches for a \emph{diff} with 
deletions between $19$ and $21$ and additions between $9$ and $11$. After the $10$ first attempts, 
the algorithm increase the step for $2$ and searches for a \emph{diff} with deletions between $18$ and $22$ and additions between 
$8$ and $10$. We originate the regular commits from the security
commits to ensure that differences in maintainability are not a consequence of
characteristics of different projects.

@@

In this research, we follow a very similar methodology to the one presented in previous work on the maintainability 
of energy-oriented fixes~\cite{cruz2019energyoriented}. As said before, the web-based source code analysis service \emph{Better Code
Hub} (BCH) is used to collect the maintainability reports of the refactorings of
each project. Table~\ref{tab:guidelines} presents the $10$ guidelines $-$ proposed
by BCH's authors for delivering software that is not difficult to
maintain~\cite{Visser:2016:OREILLY} $-$ and maps each guideline to the metric 
calculated by BCH. Theses guidelines are calculated using the metrics 
presented in~\cite{criteria:2017}. During each guideline evaluation, the tool
determines the compliance towards one guideline by establishing limits for the
percentage of code allowed to be in each of the $4$ risk severity levels
(\emph{low risk}, \emph{medium risk}, \emph{high risk}, and \emph{very high
risk}). If the project does not violate the thresholds, then it is compliant
with the guideline. These thresholds are calibrated by BCH using their own
data/experience --- using open-source and closed software systems. If a project is
compliant with a guideline, it means that it is at least $65\%$ better than the
software used by BCH to calculate the thresholds\footnote{Check the answer to
\emph{How can I adjust the threshold for passing/not passing a guideline?} at
\url{https://bettercodehub.com/docs/faq} (Accessed on \today{})}.

@@

\begin{table}[h]
\footnotesize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.66cm}L{3.3cm}L{2.7cm}}

@@

SIG defines \emph{Units} as the smallest groups of code that can be maintained
and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and
constructors in Java). One of the guidelines with which the project does not
comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}):
\emph{Write Simple Units of Code}. BCH analyzes this guideline based on the
McCabe Complexity~\cite{1702388} to calculate the number of branch points of a
method. The bar at the bottom of the figure represents the top $30$ units that
violate the guideline, sorted by severity. The violation severities are
indicated using colors, and there is a legend to help to interpret them. The
green bar represents the number of compliant branch points per unit (\emph{at
most $5$}), i.e., the number of units are compliant with ISO
$25010$~\cite{iso:2011}. Yellow, orange and red bars represent units that do not
comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint
the compliance thresholds for each severity level. If the green mark is
somewhere in the green bar it means it is compliant with a low level of
severity.

@@

The BCH tool does not compute the final score that our study needs to compare
maintainability amongst different project versions. For this, we
follow previous work on measuring the impact of energy-oriented refactorings
~\cite{cruz2019energyoriented}. Luis et al. ($2019$) proposed an equation
to capture the distance between the current state of the project and
the standard thresholds calculated by the BCH based on the insights provided in~\cite{Olivari:2018}. 
The equation considers the following:
\begin{itemize}
	\item \textbf{Size of project changes does not affect the maintainability
	difference between two versions, $\Delta M (v_{s-1},v_{s}) = M(v_{s}) - M(v_{s-1})$.} We
	aim at evaluating security patterns occurring in different projects similarly.
	Thus, the derived metric uses the \textit{raw} number of lines of code rather
  percentages for normalization purposes.
	\item \textbf{Distance to lower severity levels is less penalized than to the
	thresholds in high severity levels.} Severity level weights based on the
	severity level to count lines of code that violate maintainability guidelines.
\end{itemize}

@@

This study evaluates $1037$ security and baseline commits
from $267$ distinct open-source projects. The following section presents the
results obtained for each research question.

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.5\textwidth]{figures/impact_per_guideline.pdf}
 	\caption{Impact of the security patches for each guideline.}
	\label{fig:guidelines}
	\scalebox{0.5}{
	\begin{tabular}{r@{: }l r@{: }l}
	$WCC$ & Write Clean Code & $KACB$ & Keep Architecture Components Balanced\\
	$CACL$ & Couple Architecture Components Loosely & $SCM$ & Separate Concerns in Modules \\
	$KUIS$ & Keep Unit Interfaces Small & $WCO$ & Write Code Once \\
	$WSimpleUC$ & Write Simple Units of Code & $WShortUC$ & Write Short Units of Code
	\end{tabular}
	}
\end{figure}

Maintainability is being calculated as the mean of the guidelines computed
by BCH. Figure~\ref{fig:guidelines} shows the impact of security patches on each 
guideline individually. For projects with large 
codebases, the results retrieved for \emph{Keep Your Codebase Small} were not 
viable due to limitations of BCH. For this guideline, the tool sets a limit of 20 Person-years. However, a large amount of projects of our database surpass this limit. Thus, we decided to not consider this guideline 
in our analysis. For each type of
guideline, three types of impact of patches in the software maintainability
are presented: \emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). Next to each
type of guideline, it is shown the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test.

Regarding security commits results, we observe that maintainability decreases in
$41.76\%$ ($433$), remains the same in $19.67\%$ ($204$) and increases in $38.57\%$
($400$) of the cases after security patches are applied. The resulting
p-value of the Paired Wilcoxon signed-rank test is $4.78$x$10^{-2}$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact on the maintainability of
open-source software}.
Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $5$ guidelines: 
\emph{Write Clean Code} (p-value = $4.9$x$10^{-54}$), \emph{Keep Architecture 
Components Balanced} (p-value = $5.26$x$10^{-71}$), \emph{Couple Architecture 
Components Loosely} (p-value = $2.43$x$10^{-07}$), \emph{Keep Unit Interfaces 
Small} (p-value = $2.25$x$10^{-13}$), \emph{ Write Code Once} (p-value = 
$7.9$x$10^{-35}$). However, for each of these guidelines we also observe a substantial number of cases where the maintainability suffers a negative impact $-$ from $10\%$ to $35\%$ $-$ which highlights the need of better practices/techniques/tools to produce software. \emph{Separate Concerns in Module} 
guideline also follows the same results but without any statistical 
significance since the p-value is higher than $0.05$. According to the 
results, \emph{Write Simple Units of Code} (p-value=$0.015$ and $38.2\%$/$396$ 
of negative cases) and \emph{Write Short Units of Code} (p-value=$0.029$ and 
$38.3\%$/$397$ of negative cases) are the most affected affected guidelines.

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/maintainability_severity.pdf}
 	\caption{Maintainability difference by the vulnerabilities severity.}
	\label{fig:severity}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/maintainability_language.pdf}
  \caption{Maintainability difference by language after security patches.}
  \label{fig:lang_main}  
\end{figure}

We have also evaluated the impact on the maintainability per programming language
of open-source software, Figure~\ref{fig:lang_main}. We restrict this analysis to programming languages with at least $20$ data points, as this is a requirement for our hypothesis tests. Thus, we compare the results for \emph{Python}, \emph{C/C++}, \emph{Ruby}, \emph{PHP}, 
\emph{Objective-C/C++}, and \emph{Java}. The other languages were
integrated into \emph{Others}. The results of the Paired Wilcoxon signed-rank
test by language indicate that there is statistical evidence to support that the
maintainability of open-source software decreases for security patches for \emph{C/C++} 
(p-value = $2.5$x$10^{-5}$). 
% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.
\emph{Ruby} and \emph{PHP} languages also impact negatively the maintainability of software. Despite not having statistical significance, there are a considerable number of \emph{Ruby} and \emph{PHP} patches hindering maintainability that should not be disregarded

@@

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.4\textwidth]{figures/maintainability_cwe.pdf}
 	\caption{Maintainability difference by pattern after security patches.}
	\label{fig:pat}	
\end{figure}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}
\usepackage{footnote}
\usepackage{xfrac}
\usepackage{balance}
\usepackage{hyperref}

@@

% \title{Fixing Vulnerabilities Hinders Maintainability}
\title{\Sof{TBA - New Titles in the comments}}
%\title{Coding practices are important to produce secure software: Teach them to your students and developers!}
%\title{The reason why you should be using coding practices: safer software.}
%\title{What is the impact of security patches in software evolution?}
%\title{Hey, security engineers! You should pay more attention to your patches.}

@@

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of $10$ 
guidelines using a dataset including $1335$ security commits
\Sof{These are the initial values, should I use the final? In the end, 
we studied 1031 commits and 8 guidelines}. 
Results show 
that fixing software vulnerabilities may hinder the maintainability of 
open-source software. This is particularly the case for patches that 
address Path Traversal (CWE-22), Information Exposure (CWE-200) and Resource 
Management Errors (CWE-399). We conclude that changes to code bases to patch 
vulnerabilities need to be performed with extra care and that security-related 
code evolution should be subject in computer curricula.
\end{abstract}

@@

exploit their resources. \Sof{Update example for 2020 - outdated}As an example, in the last week of January $2019$, 

@@

details available at \url{https://thehackernews.com/2019/01/apple-facetime-privacy-hack.html} (Accessed on \today{})} --- the outcome of a design/logical
flaw that allowed callers to see and hear others without them picking up the call.\Sof{---}%\cite{cacheOut}.

@@

maintainability~\cite{4724577}, documentation to help developers adopt security 
best practices~\cite{6311252, 7927935, MESQUIDA201519} and the importance of integrating
security-related code evolution in computer curricula.

@@

  \url{https://figshare.com/s/4861207064900dfb3372}.

@@

implications; section~\ref{sec:threats} enumerates the threats to the validity of

@@

%%%%% NEEDS REVIEW
\Sof{Improve for 2nd version}
As an example, consider the patch of the TLS state-machine protocol implementation in 
OpenSSL\footnoteref{openssl} to address a vulnerability where an attacker could cause a Denial-of-Service
(DoS) attack by crafting TLS messages to trigger memory allocations
corresponding to large length values. This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2016$-$6307$ \footnote{CVE-$2016$-$6307$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2016-6307}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
evaluation. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/statem/statem.c} file\footnote{CVE-$2016$-$6307$ fix available  at
\url{https://github.com/openssl/openssl/commit/c1ef7c971d0bbf117c3c80f65b5875e2e7b024b1}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. Every SSL/TLS connection begins with a handshake which is responsible by the
negotiation between the two parties. The problem in
CVE-$2016$-$6307$ is that due to a flaw in the logic of version $1.1.0$,
memory for a message is allocated too early, prior to the excessive
message length check which can lead to a Denial of Service through memory
exhaustion if an attacker triggers up a certain message length to be allocated to service a
connection. In this case, there is only a security impact if the application does not call  \texttt{SSL\_free()} in a timely manner in the
event that the connection fails;
or, the application is working in a constrained environment where there
is very little free memory;
or, the attacker initiates multiple connection attempts such that there
are multiple connections in a state where memory has been allocated for
the connection, \texttt{SSL\_free()} has not yet been called, and there is
insufficient memory to service the multiple requests. Except in the instance of $1$ above, any Denial Of Service is likely to
be transitory because as soon as the connection fails the memory is
subsequently freed again in the \texttt{SSL\_free()} call. However there is an
increased risk during this period of application crashes due to the lack
of memory - which would then mean a more serious Denial of Service.

@@

%
Although checking the maximum size of a message seems an elementary
problem to repair, the example shows that there was a significant amount of
changes performed in the code base which produced a negative impact in the
maintainability. The developer tries to patch the
code base by introducing a new branch point to the function \texttt{read\_state\_machine}~\ref{lst:func1} which adds complexity to the software
and disrupts
one of the guidelines proposed by the Software Improvement Group (SIG) for
building maintainable software~\cite{Visser:2016:OREILLY} $-$
the \emph{Write Simple Units of Code} guideline $-$ which resulted in a negative impact on the
OpenSSL maintainability.
\Sof{-- END --}
%%%%%%% 
Software maintainability is designed as the degree to which an application is understood, 
repaired or enhanced. In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability of their
software applications. This is important because software maintainability is 
approximately $75\%$ of the cost related to a project. To answer the following three 
research questions, we use two datasets of security patches~\cite{Reis:2017:IJSSE, 10.1109/MSR.2019.00064} 
to measure the impact of
security patches on the maintainability of open-source software. 
%

\textit{\textbf{RQ1: What is the impact of security patches on the
maintainability of open-source software?}} Often security flaws require patching 
code to make software more secure.
However, \textbf{there is no evidence yet of how security patches impact the
maintainability of open-source software}. Our suspicion is that developers tend
to introduce technical debt in their software when patching the code base to
address a security flaw because they tend to choose the easiest path to solve
it. To address it, we compute the maintainability of $1335$ commits using the
\emph{Better Code Hub} tool. 
%

\textit{\textbf{RQ2: Which patterns of security patches are more likely to
affect open-source software maintainability?}}

@@

the cross-site scripting vulnerability is presented in Listing~\ref{lst:fix}. The developer added the function \texttt{htmlentities} to escape the data given by the variable

@@

\textit{\textbf{RQ3: What is the impact of security patches versus regular changes on the
maintainability of open-source software?}}

@@

We also compute the maintainability of random regular commits using the

@@

	Extract relevant data (e.g., owner and name of
	the repository, sha key of the vulnerable version, sha key of the fixed version) to the experiments and code bases from the combined dataset containing $1335$ security

@@

the outcome of mining/manually inspect $265$ GitHub projects. Reis and Abreu 

@@

dataset of $711$ test cases for $16$ different security vulnerability types. Some of the vulnerability types

@@

code used to fix the security flaw. 

@@

for references to commits involved in the patch on the National Vulnerability Database (NVD) website. However, $70\%$

@@

used to patch the vulnerability. We assumed the last commit of the patch as 

@@

In this study, we focus on computing the

@@

$1031$ security patches. $539$ of these patches are identified with Common Weakness
Enumeration IDs (CWE). In Table~\ref{tab:patterns}, the most prevailing categories are presented. To satisfy the Wilcoxon test size requirement, types of vulnerabilities
with less than $20$ instances were merged in a bigger group named \textit{Miscellaneous}.

\begin{table}[h]
\scriptsize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.1cm}L{2.1cm}L{4.3cm}}

@@

 	     	           \textbf{CWE-79} & 
Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (63 cases)& 
Lack of proper validation or escaping
 	    allows attackers to submit untrusted data to web browsers through malicious
 	    scripts that can hijack the user sessions or redirect the user to malicious
 	    sites.\\\midrule

@@

 	 	     	           \textbf{MISC} & 
Miscellaneous (666 cases) & 
This pattern comprises
 		several other security refactorings that do not have a CWE assigned and patterns that
 		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.\\
\bottomrule
\end{tabular}
\label{tab:patterns}
\end{table}

@@

The baseline dataset uses the security commits dataset as input.
Both datasets have the same size. For each
security commit, a commit with an approximate size is searched. Due to the complexity 
of some patches, it was not possible to find patches with the exact number of added and 
deleted lines. Our search uses the patch \emph{diff} to look for a new similar \emph{diff}. The algorithm integrates the following steps:

\begin{enumerate}
\item Calculates the \emph{diff} between the security patch commit (\emph{sha}) and its parent (\emph{sha-p}).
\item Selects a random commit from the project (\emph{sha-reg}).
\item Calculates the \emph{diff} between the random commit (\emph{sha-reg}) and its parent (\emph{sha-reg-p}).
\item Checks if the size from the regular change has the same/approximated size to 
the security patch.
\end{enumerate}

The pair of the random commit and its parent is accepted if the \emph{diff} 
size fits in the range size considered acceptable at the moment. The algorithm 
starts with a step of $1$. Every $10$ attempts to search for a change with the 
same \emph{diff} the step increases. Thus, if a change suffered $20$ deletions 
and $10$ additions, then, in the first $10$ attempts, the algorithm searches 
for a \emph{diff} with deletions between $19$ and $21$ and additions between 
$9$ and $11$. After the $10$ first attempts, the algorithm increase the step 
for $2$ and searches for a \emph{diff} with deletions between $18$ and $22$ 
and additions between $8$ and $10$. We originate the regular commits from the 
security commits to ensure that differences in maintainability are not a 
consequence of characteristics of different projects.

@@

In this research, we follow a very similar methodology to the one presented 
in previous work on the maintainability of energy-oriented fixes~\cite{cruz2019energyoriented}. 
As said before, the web-based source code analysis service \emph{Better Code
Hub} (BCH) is used to collect the maintainability reports of the refactorings of
each project. Table~\ref{tab:guidelines} presents the $10$ guidelines $-$ proposed
by BCH's authors for delivering software that is not difficult to
maintain~\cite{Visser:2016:OREILLY} $-$ and maps each guideline to the metric 
calculated by BCH. Theses guidelines are calculated using the metrics 
presented in~\cite{criteria:2017} and also briefly explained on Table~\ref{tab:guidelines}. 
During each guideline evaluation, the tool determines the compliance towards 
one guideline by establishing limits for the percentage of code allowed to be 
in each of the $4$ risk severity levels
(\emph{low risk}, \emph{medium risk}, \emph{high risk}, and \emph{very high
risk}). If the project does not violate the thresholds, then it is compliant
with the guideline. These thresholds are calibrated by BCH using their own
data/experience --- using open-source and closed software systems. If a project is
compliant with a guideline, it means that it is at least $65\%$ better than the
software used by BCH to calculate the thresholds\footnote{Check the answer to
\emph{How can I adjust the threshold for passing/not passing a guideline?} at
\url{https://bettercodehub.com/docs/faq} (Accessed on \today{})}.

@@

\begin{table}[h]
\scriptsize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.66cm}L{3.3cm}L{2.7cm}}

@@

\Sof{Needs improvement}SIG defines \emph{Units} as the smallest groups of code that can be maintained
and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and
constructors in Java). One of the guidelines with which the project does not
comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}):
\emph{Write Simple Units of Code}. BCH analyzes this guideline based on the
McCabe Complexity~\cite{1702388} to calculate the number of branch points of a
method. The bar at the bottom of the figure represents the top $30$ units that
violate the guideline, sorted by severity. The violation severities are
indicated using colors, and there is a legend to help to interpret them. The
green bar represents the number of compliant branch points per unit (\emph{at
most $5$}), i.e., the number of units are compliant with ISO
$25010$~\cite{iso:2011}. Yellow, orange and red bars represent units that do not
comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint
the compliance thresholds for each severity level. If the green mark is
somewhere in the green bar it means it is compliant with a low level of
severity.\Sof{--END--}

@@

The BCH tool does not compute the final score that our study needs to compare
maintainability amongst different project versions. For this, we
follow previous work on measuring the impact of energy-oriented refactorings
~\cite{cruz2019energyoriented}. Luis et al. ($2019$) proposed an equation
to capture the distance between the current state of the project and
the standard thresholds calculated by the BCH based on the insights 
provided in~\cite{Olivari:2018}. The equation provided in~\cite{cruz2019energyoriented} 
considers that the size of project changes does not affect the maintainability and
that the distance to lower severity levels is less penalized than to the
thresholds in high severity levels.
% 	thresholds in high severity levels
% 	difference between two versions
% The equation considers the following:
% \begin{itemize}
% 	\item \textbf{Size of project changes does not affect the maintainability
% 	difference between two versions, $\Delta M (v_{s-1},v_{s}) = M(v_{s}) - M(v_{s-1})$.} We
% 	aim at evaluating security patterns occurring in different projects similarly.
% 	Thus, the derived metric uses the \textit{raw} number of lines of code rather
%   percentages for normalization purposes.
% 	\item \textbf{Distance to lower severity levels is less penalized than to the
% 	thresholds in high severity levels.} Severity level weights based on the
% 	severity level to count lines of code that violate maintainability guidelines.
% \end{itemize}

@@

This study evaluates $1031$ security and baseline commits
from $265$ distinct open-source projects. The following section presents the
results obtained for each research question.

\textit{\textbf{RQ1: What is the impact of security patches on the
maintainability of open-source software?}}

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.5\textwidth]{figures/main_per_guideline.pdf}
 	\caption{Impact of the security patches per guideline.}
	\label{fig:guidelines}
	\scalebox{0.5}{
	\begin{tabular}{r@{: }l r@{: }l}
	$WCC$ & Write Clean Code & $KACB$ & Keep Architecture Components Balanced\\
	$CACL$ & Couple Architecture Components Loosely & $SCM$ & Separate Concerns in Modules \\
	$KUIS$ & Keep Unit Interfaces Small & $WCO$ & Write Code Once \\
	$WSimpleUC$ & Write Simple Units of Code & $WShortUC$ & Write Short Units of Code
	\end{tabular}
	}
\end{figure}

Maintainability is being calculated as the mean of the guidelines computed
by BCH. Figure~\ref{fig:guidelines} shows the impact of security patches on each 
guideline individually. 
For each type of
guideline, three types of impact of patches in the software maintainability
are presented: \emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). Next to each
type of guideline, it is shown the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test.

Regarding the impact of security patches for each guideline, we observe
that overall patches have a significant positive impact in $5$ guidelines: 
\emph{Write Clean Code} (p-value = $4.9$x$10^{-54}$), \emph{Keep Architecture 
Components Balanced} (p-value = $5.26$x$10^{-71}$), \emph{Couple Architecture 
Components Loosely} (p-value = $2.43$x$10^{-07}$), \emph{Keep Unit Interfaces 
Small} (p-value = $2.25$x$10^{-13}$), \emph{ Write Code Once} (p-value = 
$7.9$x$10^{-35}$). However, for each of these guidelines we also observe a 
substantial number of cases where the maintainability suffers a negative 
impact $-$ from $10\%$ to $35\%$ $-$ which highlights the need of better 
practices/techniques/tools to produce software. \emph{Separate Concerns in Module} 
guideline also follows the same results but without any statistical 
significance since the p-value is higher than $0.05$. According to the 
results, \emph{Write Simple Units of Code} (p-value=$0.015$ and $38.2\%$/$396$ 
of negative cases) and \emph{Write Short Units of Code} (p-value=$0.029$ and 
$38.3\%$/$397$ of negative cases) are the most affected affected guidelines.

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.45\textwidth]{figures/main_per_severity.pdf}
 	\caption{Maintainability difference by the vulnerabilities severity.}
	\label{fig:severity}
\end{figure}

For projects with large 
codebases, the results retrieved for \emph{Keep Your Codebase Small} were not 
viable because they retrieved values above the limit set by BCH ($20$ Person-years).
The \emph{Automated Tests} guideline was also not considered because the BCH does not return
the amount of lines of code that complies with the guideline -- which is one of the variables
our formula uses to calculate the maintainability of the project regarding each guideline. 
Thus, we decided to not consider both guidelines in our analysis.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.5\textwidth]{figures/main_per_language.pdf}
  \caption{Maintainability difference by language after security patches.}
  \label{fig:lang_main}  
\end{figure}

We have also evaluated the impact on the maintainability per programming language
of open-source software, Figure~\ref{fig:lang_main}. We restrict this analysis to 
programming languages with at least $20$ data points, as this is a requirement 
for our hypothesis tests. Thus, we compare the results for \emph{Python}, \emph{C/C++}, \emph{Ruby}, \emph{PHP}, 
\emph{Objective-C/C++}, and \emph{Java}. The other languages were
integrated into \emph{Others}. The results of the Paired Wilcoxon signed-rank
test by language indicate that there is statistical evidence to support that the
maintainability of open-source software decreases for security patches for \emph{C/C++} 
(p-value = $2.5$x$10^{-5}$). 
% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.
\emph{Ruby} and \emph{PHP} languages also impact negatively the maintainability of software. 
Despite not having statistical significance, there are a considerable number of \emph{Ruby} 
and \emph{PHP} patches hindering maintainability that should not be disregarded.

@@

Regarding security commits results, we observe that maintainability decreases in
$41.61\%$ ($429$), remains the same in $19.69\%$ ($203$) and increases in $38.70\%$
($399$) of the cases after security patches are applied. The resulting
p-value of the Paired Wilcoxon signed-rank test is $4.9$x$10^{-2}$.
Since the p-value is below the significance level of $0.05$, we argue that
\emph{security patches have a negative impact on the maintainability of
open-source software}.

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.4\textwidth]{figures/main_per_cwe.pdf}
 	\caption{Maintainability difference by pattern after security patches.}
	\label{fig:pat}	
\end{figure}
